Surface Source Extension, Reflection Case

### 2019-12-17 repro/papers/sserefl/paper.tex
more edits

Now assume that $\kappa=\kappa_0 \in \bR^+, \rho=\rho_0 \in \bR^+, c_0
= \sqrt{\kappa_0/rho_0}$ are constant. 
In $z \ne z_s$, the right hand side 
vanishes, so the solution must be locally a combination of plane
waves; causality implies that
\[
p(z,t)=a\left(t \mp \frac{z-z_s}{c_0}\right), \, v(z,t)=\pm b\left(t \mp
  \frac{z_s}{c_0}\right)
\]
From the second dynamical equation (Newton's law) it follows that $b =
a/(\rho_0 c_0)$. The singularity on the LHS of the first dynamical
equation (constitutive law) is
\[
\kappa [v]_{z=z_s}\delta(z-z_s) =
2\frac{\kappa}{c_0}b\delta(z-z_s) = 2a\delta(z-z_s).
\] 
This must in turn equal the RHS of the constitutive law, whence
$a=w/2$. Thus
\begin{eqnarray}
\label{eqn:sol1dptsrc}
p(z,t) &=& \frac{1}{2}w\left(t - \frac{|z-z_s|}{c_0}\right) \nonumber \\
v(z,t) &=& \mbox{sgn}(z-z_s)\frac{1}{2\rho_0 c_0}w\left(t -\frac{|z-z_s|}{c_0}\right)
           \nonumber \\
\end{eqnarray}
This result (computation of the Green's function for the acoustic
system) permits an explicit expression for the system with a
space-time source:
\begin{eqnarray}
\label{eqn:awe1d}
\frac{\partial p}{\partial t} +\kappa_0\frac{\partial 
  v}{\partial z} &=& f(z,t) \nonumber\\
\rho_0 \frac{\partial v}{\partial t} + \frac{\partial p}{\partial 
  z}&=&0\nonumber\\
 p,v&=&0, t \ll 0. 
\end{eqnarray}
Since
\[
  f(z,t) = \int dz_s\,f(z_s,t)\delta(z-z_s)
\]
obtain
\begin{eqnarray}
\label{eqn:sol1d}
p(z,t) &=& \frac{1}{2}\int dz_s f\left(z_s,t - \frac{|z-z_s|}{c_0}\right) \nonumber \\
v(z,t) &=& \frac{1}{2\rho_0 c_0} \int dz_s \mbox{sgn} (z-z_s) f\left(z_s,t - \frac{|z-z_s|}{c_0}\right)
           \nonumber \\
\end{eqnarray}

@@

The key is to understand the perturbational impulse response. That is,
set $w(t)=H(t)$. Then the identity \ref{eqn:sol1d} implies that the
inhomogeneous term in the first equation of the system \ref{eqn:dawe1d}
is 
\begin{equation}
  \label{eqn:pertrhs}
-\delta \kappa(z) \frac{\partial v}{\partial z}(z,t) = \frac{\delta
  \kappa(z_s)}{\rho c}H(t)\delta(z-z_s) +
\frac{\delta\kappa(z)}{2\kappa}\delta\left(t - \frac{|z-z_s|}{c}\right)
\end{equation}
Correspondingly, the solution $(\delta p, \delta v)$ is the sum of
three components:
\begin{eqnarray}
  \label{eqn:pertrhsterms}
  \delta p & = & \delta p_s + \delta p_d \nonumber\\
  \delta v & = & \delta v_s + \delta v_d 
\end{eqnarray}
The first component satisfies
\begin{eqnarray}
  \label{eqn:dawe1ds}
  \frac{\partial \delta p_s}{\partial t} +\kappa\frac{\partial
  \delta v_s}{\partial z}   &=&  \frac{\delta
  \kappa(z_s)}{\rho c}H(t)\delta(z-z_s)\nonumber\\
  \rho \frac{\partial \delta v_s}{\partial t} + \frac{\partial \delta p_s}{\partial
  z}&=&0\nonumber\\
  \delta p_s,\delta v_s&=&0, t < 0.
\end{eqnarray}
The initial conditions are free to be chosen, and the choice above is convenient.

Note that \ref{eqn:dawe1ds} is precisely a system of the form
\ref{eqn:awe1d}, so from the formulae \ref{eqn:sol1d} with $w(t)=\frac{\delta
  \kappa(z_s)}{\rho c}H(t)$
\begin{eqnarray}
\label{eqn:dsol1ds}
\delta p_s(z,t) &=& \frac{\delta
  \kappa(z_s)}{2\rho c}H\left(t - \frac{|z-z_s|}{c}\right) \nonumber \\
\delta v_s(z,t) &=& \pm\frac{\delta \kappa(z_s)}{2\rho^2 c^2}H\left(t - \frac{|z-z_s|}{c}\right)
\end{eqnarray}

Define $P_d(z,t;\zeta), V_d(z,t;\zeta)$ to be the solution of the system
\begin{eqnarray}
  \label{eqn:dawe1dd}
  \frac{\partial P_d}{\partial t} +\kappa\frac{\partial
  V_d}{\partial z}   &=&   \frac{\delta\kappa(\zeta)}{2\kappa}\delta(z-\zeta)\delta\left(t - \frac{|\zeta-z_s|}{c}\right)\nonumber\\
  \rho \frac{\partial V_d}{\partial t} + \frac{\partial P_d}{\partial
  z}&=&0\nonumber\\
  P_d,V_d&=&0, t < 0.
\end{eqnarray}
Then
\begin{eqnarray}
  \label{eqn:intrep1}
  \delta p_d(z,t) & = & \int d\zeta\, P_d(z,t;\zeta)\\
  \delta v_d(z,t) & = & \int d\zeta\, V_d(z,t;\zeta)
\end{eqnarray}

The system \ref{eqn:dawe1dd} is precisely of the form
\ref{eqn:awe1d} with $z_s$ replaced by $\zeta$ and
\[
  w(t)= \frac{\delta\kappa(\zeta)}{2\kappa}\delta\left(t - \frac{|\zeta-z_s|}{c}\right)
\]
Thus from \ref{eqn:sol1d}
\begin{eqnarray*}
  P_d(z,t;\zeta) &=& \frac{\delta\kappa(\zeta)}{4\kappa} \delta\left(t - \frac{|z-\zeta|+|\zeta-z_s|}{c}\right) \nonumber \\
  V_d(z,t;\zeta) &=& \pm\frac{\delta\kappa(\zeta)}{4\kappa \rho c} \delta\left(t - \frac{|z-\zeta|+|\zeta-z_s|}{c}\right) 
\end{eqnarray*}

\begin{equation}
\label{eqn:imp1d}
D(S[\kappa]H)\delta \kappa (t) = \delta p(0,t)=  \frac{\delta 
  \kappa\left(\frac{ct}{2}\right)}{8\kappa} 
\end{equation}
whence for arbitrary causal wavelet $w$,
\begin{equation}
\label{eqn:fwd1d}
D(S[\kappa]w)\delta \kappa (t) = \int_0^t\,ds\,w'(t-s)\frac{\delta 
  \kappa\left(\frac{cs}{2}\right)}{8\kappa} 
\end{equation}

\append{Plane Waves}
Re-write the plane wave system \ref{eqn:awepw} as
\begin{eqnarray}
\label{eqn:awepw1}
\frac{\partial \tp}{\partial t} + \tk\frac{\partial
  \tv}{\partial z} &=& (1-c_0^2\xi^2)^{-1}\tf \nonumber\\
\rho \frac{\partial \tv}{\partial t} + \frac{\partial \tp}{\partial
  z}&=&0\nonumber\\
\tp,\tv&=&0, t \ll 0,
\end{eqnarray}
in which $\tk(\xi,z)=(1-c(z)^2\xi^2)^{-1}\kappa(z)$ is a scaled version of
bulk modulus. Define the vertical plane wave velocity as
\[
\tc(\xi,z)=\frac{c(z)}{1-c(z)^2\xi^2} =
\left(\frac{1}{c(z)^2}-\xi^2\right)^{-1/2}.
\]
Comparing the plane wave system \ref{eqn:awepw1} with the 1D system
\ref{eqn:awe1d}, and using the 1D solution \ref{eqn:sol1d}, obtain
\begin{eqnarray}
\label{eqn:solpw}
\tp(\xi,z,t) &=& \frac{1}{2}(1-c_0^2\xi^2)^{-1}\int dz_s \tf\left(\xi,z_s,t - \frac{|z-z_s|}{\tc_0(\xi)}\right) \nonumber \\
\tv(\xi,z,t) &=& \frac{1}{2\rho_0 c_0}(1-c_0^2\xi^2)^{-1} \int dz_s \mbox{sgn} (z-z_s)\tf\left(\xi,z_s,t - \frac{|z-z_s|}{\tc_0(\xi)}\right)
           \nonumber \\
\end{eqnarray}

>>>>

Now assume that $\kappa=\kappa_0 \in \bR^+, \rho=\rho_0 \in \bR^+, c_0
= \sqrt{\kappa_0/rho_0}$ are constant. 
In $z \ne z_s$, the right hand side 
vanishes, so the solution must be locally a combination of plane
waves; causality implies that
\[
p(z,t)=a\left(t \mp \frac{z-z_s}{c_0}\right), \, v(z,t)=\pm b\left(t \mp
  \frac{z_s}{c_0}\right)
\]
From the second dynamical equation (Newton's law) it follows that $b =
a/(\rho_0 c_0)$. The singularity on the LHS of the first dynamical
equation (constitutive law) is
\[
\kappa [v]_{z=z_s}\delta(z-z_s) =
2\frac{\kappa}{c_0}b\delta(z-z_s) = 2a\delta(z-z_s).
\] 
This must in turn equal the RHS of the constitutive law, whence
$a=w/2$. Thus
\begin{eqnarray}
\label{eqn:sol1dptsrc}
p(z,t) &=& \frac{1}{2}w\left(t - \frac{|z-z_s|}{c_0}\right) \nonumber \\
v(z,t) &=& \mbox{sgn}(z-z_s)\frac{1}{2\rho_0 c_0}w\left(t -\frac{|z-z_s|}{c_0}\right)
           \nonumber \\
\end{eqnarray}
This result (computation of the Green's function for the acoustic
system) permits an explicit expression for the system with a
space-time source:
\begin{eqnarray}
\label{eqn:awe1d}
\frac{\partial p}{\partial t} +\kappa_0\frac{\partial 
  v}{\partial z} &=& f(z,t) \nonumber\\
\rho_0 \frac{\partial v}{\partial t} + \frac{\partial p}{\partial 
  z}&=&0\nonumber\\
 p,v&=&0, t \ll 0. 
\end{eqnarray}
Since
\[
  f(z,t) = \int dz_1\,f(z_s,t)\delta(z-z_1)
\]
obtain
\begin{eqnarray}
\label{eqn:sol1d}
p(z,t) &=& \frac{1}{2}\int dz_1 f\left(z_1,t - \frac{|z-z_1|}{c_0}\right) \\
v(z,t) &=& \frac{1}{2\rho_0 c_0} \int dz_1 \mbox{sgn} (z-z_1) f\left(z_1,t - \frac{|z-z_1|}{c_0}\right)\\
\end{eqnarray}

@@

From the second equation in the pair \ref{eqn:sol1d}
\[
  \frac{\partial v}{\partial z}(z,t) = \frac{1}{\rho_0 c_0} f(z,t) 
\]
\begin{equation}
  \label{eqn:sol1ddvdz}
 - \frac{1}{2\kappa_0} \int\, dz_2\, \frac{\partial f}{\partial t}\left(z_2,t - \frac{|z-z_2|}{c_0}\right)
\end{equation}
Thus using the first equations in \ref{eqn:sol1d} and \ref{eqn:dawe1d}, 
\[
  \delta p(z,t) =- \frac{1}{2} \int \, dz_1 \, \delta
  \kappa(z_1)\frac{\partial v}{\partial z}\left(z_1,t - \frac{|z-z_1|}{c_0}\right)
\]
\[
  =-\frac{1}{2\rho_0c_0} \int \,dz_1\, \delta
  \kappa(z_1) f \left(z_1,t - \frac{|z-z_1|}{c_0}\right)
  \]
\begin{equation}
  \label{eqn:sol1ddp}
  + \frac{1}{4\kappa_0} \int\,dz_2 \, \int \,dz_1\,\delta 
  \kappa(z_1) \frac{\partial f}{\partial t}\left(z_2,t - \frac{|z_2-z_1|+|z_1-z|}{c_0}\right)
\end{equation}

An important special case is the 1D point source,
$f(z,t)=w(t)\delta(z-z_s)$: then
\[
  \delta p(z,t) = -\frac{\delta \kappa(z_s)}{2\rho_0c_0}w\left(t - \frac{|z-z_s|}{c_0}\right)
\]
\begin{equation}
  \label{eqn:sol1ddppw}
  +\frac{1}{4\kappa_0} \int \,dz_1\,\delta 
  \kappa(z_1) \frac{\partial w}{\partial t}\left(t - \frac{|z_s-z_1|+|z_1-z|}{c_0}\right)
\end{equation}

\append{Plane Waves}
Re-write the plane wave system \ref{eqn:awepw} as
\begin{eqnarray}
\label{eqn:awepw1}
\frac{\partial \tp}{\partial t} + \tk\frac{\partial
  \tv}{\partial z} &=& (1-c_0^2\xi^2)^{-1}\tf \nonumber\\
\rho \frac{\partial \tv}{\partial t} + \frac{\partial \tp}{\partial
  z}&=&0\nonumber\\
\tp,\tv&=&0, t \ll 0,
\end{eqnarray}
in which $\tk(\xi,z)=(1-c(z)^2\xi^2)^{-1}\kappa(z)$ is a scaled version of
bulk modulus. Define the vertical plane wave velocity as
\[
\tc(\xi,z)=\frac{c(z)}{\sqrt{1-c(z)^2\xi^2}} =
\left(\frac{1}{c(z)^2}-\xi^2\right)^{-1/2}.
\]
Comparing the plane wave system \ref{eqn:awepw1} with the 1D system
\ref{eqn:awe1d}, and using the 1D solution \ref{eqn:sol1d}, obtain
\begin{eqnarray}
\label{eqn:solpw}
\tp(\xi,z,t) &=& \frac{1}{2}(1-c_0^2\xi^2)^{-1}\int dz_s \tf\left(\xi,z_s,t - \frac{|z-z_s|}{\tc_0(\xi)}\right) \nonumber \\
\tv(\xi,z,t) &=& \frac{1}{2\rho_0 c_0}(1-c_0^2\xi^2)^{-1} \int dz_s \mbox{sgn} (z-z_s)\tf\left(\xi,z_s,t - \frac{|z-z_s|}{\tc_0(\xi)}\right)
           \nonumber \\
\end{eqnarray}
